---
title: "Data Wrangling in TidyR"
author: "Ruth Enriquez"
date: "2022-08-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)
library(tidyverse)
library(here)
library(janitor)
```

## Read in the WB data files

```{r}
wb_indicators <- read_csv(here::here("data", "wb_indicators.csv"), na = c("..", ""))
wb_metadata <- read_csv(here("data", "wb_indicators_metadata.csv"))
```

### Making the `wb_indicators` table tidy, using `pivot_longer()`

### pivot_longer() to get years into a single column

```{r}
wb_indicators_long <- wb_indicators |> 
  pivot_longer(cols = `2001 [YR2001]`:`2020 [YR2020]`,
               names_to = "year",
               values_to = "indicator_value")
```

### Separate to clean up the year column

### select is the column one, filter is the row one

### use the unique() it'll tell you all the different unique levels

### drop_na() does complete row's where it will be dropped

### Widen the variables currently trapped in `Series Name` to be spread across multiple columns, and get the values that populate the columns from `indicator value`

```{r}
wb_clean <- wb_indicators_long |> 
  separate(col = year, into = c("year", "year_chr"), sep = " ") |> 
  select(-year_chr, -`Country Code`, -`Series Code`) |> 
  mutate(year = as.numeric(year)) |> 
  drop_na('Series Name') |> 
  pivot_wider(names_from = 'Series Name', values_from = indicator_value)

names(wb_clean) <- c("country", "year", "access_clean_fuels_pp", "access_electricity_pp", "co2_emissions_kt", "fossil_fuel_cons_pct", "water_stress")

wb_subset <- wb_clean |> 
  filter(country %in% c("Algeria", "Barbados", "Bulgaria", "Chile"))

ggplot(data = wb_subset, aes(x = year, y = co2_emissions_kt)) +
  geom_line(aes(color = country)) +
  facet_wrap(~country)
#if you wanted to exclude stuff
#exclude_this <- c("Algeria", "Barbados", "Bulgaria", "Chile")
#wb_example  <- 
```

## filtering rows

Example 1: Filter to only keep exact matches

```{r}
ex_1 <- starwars |> 
  filter(height > 180)
```

```{r}
ex_2 <- starwars |> 
  filter(eye_color == "blue")
```

```{r}
ex_3 <- starwars |> 
  filter(homeworld == "Naboo")
```

Examples: filtering to satisfy multiple conditions

```{r}
ex_4 <- starwars |> 
  filter(height > 180 & homeworld == "Tatooine")
```


```{r}
ex_5 <- starwars |> 
  filter(hair_color == "brown" & species == "Human")
```

### Filtering OR statements, a row will get returned if ANY of your statements are true

```{r}
ex_6 <- starwars |> 
  filter(height > 180 | eye_color == "yellow")
```

create a subset that keeps rows if homeworld is Endor or species is droid

```{r}
ex_7 <- starwars |> 
  filter(homeworld == "Endor" | species == "Droid")
```

# OR statement using picking variables from same column

```{r}
ex_8 <- starwars |> 
  filter(eye_color == "blue" | eye_color == "brown" | eye_color == "red")

#More efficient OR statement
#%in% every individual row, it will check i.e. row 1? blue? no, brown? no, red? no, next! VERSUS == it will cycle through the vectors given i.e. row 1 blue? next row 2 brown?
ex_9 <- starwars |> 
  filter(eye_color %in% c("blue", "brown", "red"))

#THIS IS NEVER WHAT YOU WANT TO DO
ex_10 <- starwars |> 
  filter(eye_color == c("blue", "brown", "red"))
#create a subset homeworld is Endor OR Tatooine or Naboo AND height <150 "," can also be seen as an &
ex_11 <- starwars |> 
  filter(homeworld %in% c("Endor", "Tatooine", "Naboo")
         & height < 150)
```

Examples with the negate !

```{r}
ex_12 <- starwars |> 
  filter(species != "Human")

ex_13 <- starwars |> 
  filter(!species %in% c("Human", "Ewok", "Wookiee"))
```

## `select()`: select or exclude columns

```{r}
ex_14 <- wb_clean |> 
  select(country, year, co2_emissions_kt)

ex_15 <- wb_clean |> 
  select(year:fossil_fuel_cons_pct)

ex_16 <- wb_clean |> 
  select(year:fossil_fuel_cons_pct, -access_electricity_pp)
```

```{r}
ex_17 <- wb_clean |> 
  filter(country %in% c("Bolivia", "Chile", "Mexico")) |> 
  select(country:fossil_fuel_cons_pct)
```

```{r}
ex_18 <- wb_clean |> 
  select(year_new = year,
         emissions = co2_emissions_kt,
         country_name = country)

#Select only columns for country, year, and water stress, and reorder them as year, country water stress THEN filter to only include observations for any three countries of your choosing

ex_19 <- wb_clean |> 
  select(year, country, water_stress) |> 
  filter(country %in% c("Australia", "Aruba", "Lebanon"))
```

