---
title: "Data Wrangling in TidyR"
author: "Ruth Enriquez"
date: "2022-08-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)
library(tidyverse)
library(here)
library(janitor)
```

## Read in the WB data files

```{r}
wb_indicators <- read_csv(here::here("data", "wb_indicators.csv"), na = c("..", ""))
wb_metadata <- read_csv(here("data", "wb_indicators_metadata.csv"))
```

### Making the `wb_indicators` table tidy, using `pivot_longer()`

### pivot_longer() to get years into a single column

```{r}
wb_indicators_long <- wb_indicators |> 
  pivot_longer(cols = `2001 [YR2001]`:`2020 [YR2020]`,
               names_to = "year",
               values_to = "indicator_value")
```

### Separate to clean up the year column

### select is the column one, filter is the row one

### use the unique() it'll tell you all the different unique levels

### drop_na() does complete row's where it will be dropped

### Widen the variables currently trapped in `Series Name` to be spread across multiple columns, and get the values that populate the columns from `indicator value`

```{r}
wb_clean <- wb_indicators_long |> 
  separate(col = year, into = c("year", "year_chr"), sep = " ") |> 
  select(-year_chr, -`Country Code`, -`Series Code`) |> 
  mutate(year = as.numeric(year)) |> 
  drop_na('Series Name') |> 
  pivot_wider(names_from = 'Series Name', values_from = indicator_value)

names(wb_clean) <- c("country", "year", "access_clean_fuels_pp", "access_electricity_pp", "co2_emissions_kt", "fossil_fuel_cons_pct", "water_stress")

wb_subset <- wb_clean |> 
  filter(country %in% c("Algeria", "Barbados", "Bulgaria", "Chile"))

ggplot(data = wb_subset, aes(x = year, y = co2_emissions_kt)) +
  geom_line(aes(color = country)) +
  facet_wrap(~country)
#if you wanted to exclude stuff
#exclude_this <- c("Algeria", "Barbados", "Bulgaria", "Chile")
#wb_example  <- 
```
